---
title: "Reproducible Research Assignment 1"
author: "Kristina Elliott"
output: 
    html_document
--
##Loading the data
First you will want to download the dataset and place into a folder on your computer. Then set your work directory to folder where you placed the dataset.
I am working in Window and in RStudio. 

Save the data into a variable. 
```{r}
ActivityData <- read.csv("activity.csv")
head(ActivityData)
```

##What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day while ignore the missing values.
```{r}
TotalNumSteps <- aggregate(steps~date, ActivityData, sum)
head(TotalNumSteps)
```
2. Make a histogram of the toal number of steps taken each day. 
```{r}
hist(TotalNumSteps$steps, main = "Total Number of Steps per Day", xlab = "Tolal Number of Steps", ylab = "Frequency", breaks = 10, col = "green")
```

3. Calculate and report the mean and median of the total number of steps taken per day.
```{r}
MeanTotalSteps <- as.integer(mean(TotalNumSteps$steps, na.rm = TRUE))
MedianTotalSteps <- median(TotalNumSteps$steps, na.rm = TRUE)
```
The mean of the total number of steps taken per day is `r MeanTotalSteps`. The median of the total number of steps taken per day is `r MedianTotalSteps`.

##What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.
```{r}
Interval <- aggregate(steps~interval, ActivityData, mean)
plot(Interval$interval, Interval$steps, type = "l", xlab = "Interval (calculated every 5 minutes)", ylab = "Number of Steps", main = "Average Number of Steps Taken per Day in 5 Minute Intervals")
```

2. Which 5 minute interaval, on average across all the days in the data set contains the maximum number of steps?
```{r}
MaxInterval <- Interval[which.max(Interval$steps),1]
```
The interaval containing the maximum number of steps is `r MaxInterval`.

##Imputing missing values
1. Calculate and report the total number of missing values in the dataset.
```{r}
NumofNA <- sum(is.na(ActivityData$steps))
```
The total number of missing values in the dataset is `r NumofNA`.

2. Devise a strategy for filling in all of the missing values in the dataset. 

Find the mean of the data with the same interval to substute in for missing values.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
ActivityDataNoNA <- ActivityData
NAValues <- is.na(ActivityDataNoNA$steps)
average_interval <- tapply(ActivityDataNoNA$steps, ActivityDataNoNA$interval, mean, na.rm=TRUE, simplify = TRUE)
ActivityDataNoNA$steps[NAValues] <- average_interval[as.character(ActivityDataNoNA$interval[NAValues])]
```

4. Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.
```{r}
TotalNumStepsNoNA <- aggregate(steps~date, ActivityDataNoNA, sum)

hist(TotalNumStepsNoNA$steps, main = "Total Number of Steps per Day including Missing Values", xlab = "Tolal Number of Steps", ylab = "Frequency", breaks = 10, col = "green")

MeanTotalStepsNoNA <- as.integer(mean(TotalNumStepsNoNA$steps, na.rm = TRUE))

MedianTotalStepsNoNA <- as.integer(median(TotalNumStepsNoNA$steps, na.rm = TRUE))
```

The mean of the total number of steps taken per day is `r MeanTotalStepsNoNA`. The median of the total number of steps taken per day is `r MedianTotalStepsNoNA`.

Do these values differ from the estimates from the first part of the assignment? 
   
    Mean with missing values : `r MeanTotalSteps` 
    Mean with No missing values: `r MeanTotalStepsNoNA`
    Median with missing values : `r MedianTotalSteps` 
    Median with No missing values: `r MedianTotalStepsNoNA`    

What is the impact of imputing missing data on the estimates of the total daily number of steps? 
    There was not much impact from the missing data.



##Are the differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
weekend <- c("Saturday", "Sunday")
ActivityDataNoNA$DW <- as.factor(ifelse(is.element(weekdays(as.Date(ActivityDataNoNA$date)), weekend), "Weekend", "Weekday"))
```
2. Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r}
IntervalNADW <- aggregate(steps~interval + DW, ActivityDataNoNA, mean)
library(lattice)
xyplot(IntervalNADW$steps~IntervalNADW$interval|IntervalNADW$DW,main = "Average Number of Steps per Intervals by Type of Day", xlab = "Interval", ylab = "Steps", layout = c(1,2), type = "l")
```
